\documentclass{article}
\usepackage{proof}
\usepackage{bussproofs}
\usepackage{xcolor}
\usepackage{url}
\usepackage{framed}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{mathtools}

\begin{document}
\title{Algorithm for Model-Based Quantifier Instantiation}
\author{Arjun Viswanathan}
\date{}
\maketitle

\section{Algorithm}
	\begin{enumerate}
		\item Given formula $F$, convert it to negation 
		normal form and apply Skolemization to obtain an 
		equisatisfiable formula of the form\\
		$\phi \land \forall x_1, ..., x_n. \psi [x_1, ..., x_n]$.
		\item Is $\phi$ satisfiable?
		\begin{itemize}
			\item No. Then, \\ 
			$\phi \land \forall x_1, ..., x_n. \psi [x_1, ..., x_n]$ is unsatisfiable.\\
			Return \textbf{UNSAT}.
			\item Yes, with model $M$. Then, \\
			Does $M$ satisfy $\forall x_1, ..., x_n. \psi [x_1, ..., x_n]$?\\
			or\\
			Is $\neg\hat{\psi}$ satisfiable?\\
			where $\neg\hat{\psi} = \psi$ with values of $M$
			substituted.
			\begin{itemize}
				\item No. Then, \\
				$\hat{\psi}$ is valid, that is, \\
				$\forall x_1, ..., x_n. \psi [x_1, ..., x_n]$ is satisfiable with model $M$.\\
				Return \textbf{SAT} with model \textbf{M}.
				\item Yes, by model $\neg\hat{\psi}[v_1, ..., v_n]$. Then,\\ 
				$M$ isn't a model of 
				$\forall x_1, ..., x_n. \psi [x_1, ..., x_n]$
				by countermodel $v_1, ..., v_n$.\\
				Let $\phi := \phi \land \psi[v_1, ..., v_n]$
				and restart step $2$.
			\end{itemize}
		\end{itemize}
	\end{enumerate}

\section{Example}
	\begin{itemize}
		\item $F :\ 0 \le a < 3 \land \forall x . x + a >_u 1$.
		\item Is $0 \le a < 3$ satisfiable? \\
			\null\qquad Yes, by $M = \{a := 0\}$.
		\item Is $\neg x + 0 >_u 1$ satisfiable? \\
			\null\qquad Yes by $\{x := 0\}$.
		\item Now,\\
			$F :\ 0 \le a < 3 \land 0 + a >_u 1 \land \forall x . x + a >_u 1$.
		\item Is $0 \le a < 3 \land 0 + a >_u 1$ 
		satisfiable?\\
			\null\qquad Yes, by $M = \{a := 2\}$.
		\item Is $\neg x + 2 >_u 1$ satisfiable? \\
		\null\qquad No.\\
		\null\qquad Thus, the original formula $F$ is 
		satisfiable by $M = \{a := 2\}$.
	\end{itemize}
\section{Fully Quantified Formulas}
	A quantified formula $\forall x_1, ..., x_n. \psi [x_1, ..., x_n]$ is satisfiable iff 
	$\neg \psi [x_1, ..., x_n]$ is unsatisfiable.\\
	Otherwise, it is unsatisfiable. So if there is no 
	quantifier-free part in $F$, the above procedure 
	reduces to expanding quantifiers with all bit-vector
	values of the corresponding bit-width.
	
\section{Example}
	Consider $F$.
	\begin{itemize}
		\item $F :\ \forall x.\ x \ge_u 0$
		\item Since the quantifier-free part of 
		$F$ is just $\top$, it doesn't place any 
		restrictions on the model $M$. 
		\item So, our problem reduces to checking whether\\
		$\neg (x \ge_u 0)$ is unsatisfiable.\\
		It happens to be unsatisfiable and $F$ 
		is satisfiable.
		\item The only way to do this for some formulas
		would be to instantiate $x$ with all possible
		values which would be exponential in the size 
		of $x$. In many cases, CVC4 has rewrite rules, 
		and without those, invertibility conditions 
		reduce this work.
	\end{itemize}
	Now consider $F$.
	\begin{itemize}
		\item $F :\ \forall x.\ x \ge_u 1$
		\item Since the quantifier-free part of 
		$F$ is just $\top$, it doesn't place any 
		restrictions on the model $M$. 
		\item So, our problem reduces to checking whether\\
		$\neg (x \ge_u 1)$ is unsatisfiable.
		\item The SMT solver might try model 
		$M = \{x := 0\}$ but \\
		$\neg (0 \ge_u 1)$ is not satisfiable.
		\item The solver then tries model 
		$M = \{x := 1\}$.\\
		$\neg (1 \ge_u 1)$ is satisfiable
		and thus $F$ is unsatisfiable.
		\item Assuming the SMT solver tries values 
		in increasing order starting from $0$, this 
		could turn into a cumbersome process, if we were
		checking say, $\forall x.\ x \ge_u 2^{64}-1$ for
		64-bit $x$.
\end{itemize}
\end{document}